maxent model avec xvalidation (year-based test training)

### 2019-04-17 src/paper2_maxent_avec_xvalidation.Rmd
finally finished the code to go through the folds, get aic and put into a simple spreadsheet for analysis

```{r train and test inc AIC for reg val}
modelmonth <- cbind(paste0(mthna ,unique(prback$year)))
aicetc <- read.csv("../data/env/AICcCalculations.csv", header = TRUE) #pre-created csv file
betavallist <- cbind(aicetc$regularization)
nfolds <- 5

for (year in 1:nrow(modelmonth)){
  timeslice <- modelmonth[year] #testing timeslice
  yr <- str_sub(timeslice, -4) #the year being used for testing
  train <- subset(prback, year == yr) #training data is all years but testing
  trainpa <- train$occurrence #creates a vector of presence/background ID (1/0)
  train  <- subset(train, select = -c(year, occurrence)) #don't need year and occurrence for modelling

  for (h in 1:nrow(betavallist)){
    betamval <-  betavallist[h]
    mod <- maxent(x = train, p = trainpa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("replicates=", nfolds), "replicatetype=crossvalidate", paste0("betamultiplier=", betavallist[h])), path = paste0("C:/Users/bunny/Documents/GitHub/Chapter_1/output/maxent/p2/models/", timeslice)) #here the full path is needed

@@

    #count number of non-zero coefficients, excluding the final four rows
    lambdas_list <- list.files(paste0("../output/maxent/p2/models/", timeslice), pattern = "*lambdas", full.names = TRUE)
    samppred_list <- list.files(paste0("../output/maxent/p2/models/", timeslice), pattern = "*samplePredictions.csv", full.names = TRUE)
    for(fold in 1:nfolds){
      fn <- fold
      modn <- fn-1
      lamf <- lambdas_list[fold]
      sampf <- samppred_list[fold]
      coefile <- read.csv(paste0(lamf), header = FALSE, stringsAsFactors=FALSE)
      coefile <- head(coefile, -4)
      coefile2 <- subset(coefile, V2 != 0)
      nocoeff <- nrow(coefile2)
      bick <- sum(as.numeric(coefile2[, 2])) #dodgy 
      bic <- (bick*log(sum(trainpa == 1)) - 2*11) #dodgy
      #natural log
      samplepred <- read.csv(paste0(sampf), header = TRUE)
      samplepred$nat_log <- log(samplepred$Raw.prediction)
      sumsamplepred <- sum(samplepred$nat_log)

@@

@@

      notrainingpoints <- maxresults$X.Training.samples[fold]
      testauc <- maxresults$Test.AUC[fold]
      trainauc <- maxresults$Training.AUC[fold]
      modno <- maxresults$Species[fold]

   #test auc scores and other lovely things you might want to include
   # ev <- evaluate(p = testp, a= testb, model = mod)
    #evauc <- ev@auc
    #evcor <- ev@cor
    #evpcor <- ev@pcor
    #evnp <- ev@np
    #evna <- ev@na
    #evmaxtprtnr <- threshold(ev, 'spec_sens')
    #evmaxkap <- threshold(ev, 'kappa')
    #evno_omission <- threshold(ev, 'no_omission')
    #evprevalence <- threshold(ev, 'prevalence')
    #evequal_sens_spec <- threshold(ev, 'equal_sens_spec')
    #evsensitivity <- threshold(ev, 'sensitivity')

      aicetc$train_AUC <- trainauc
      aicetc$test_AIC <- testauc
      aicetc$no_trainng_points <- notrainingpoints
      aicetc$test_coefficients <- nocoeff
      aicetc$test_sum.of.logs <- sumsamplepred
      #aicetc$test_AUC <- evauc
      #aicetc$test_evcor <- evcor
      # aicetc$test_evpcor <- evpcor
      # aicetc$test_nopresence <- evnp
      # aicetc$test_nobackground <- evna
      # aicetc$test_maxsumsenspec <- evmaxtprtnr
      # aicetc$test_maxkap <- evmaxkap
      # aicetc$test_no_omission <- evno_omission
      # aicetc$test_prevalence <- evprevalence
      # aicetc$test_equal_sens_spec <- evequal_sens_spec
      # aicetc$test_sensitivity <- evsensitivity
      aicetc$test_bic <- bic #not sure this is correct...
      aicetc$beta <- betaval

@@

    #calculate minimal predicted area using ecospat
    aicetc$mpa_0.9_cum <- ecospat.mpa(samplepred$Cumulative.prediction, perc = 0.9) #90%
    }
    aicetc$AIC <- ((2*aicetc$coefficients)-(2*aicetc$sum.of.logs))
aicetc$AICc <- (-2*(aicetc$sum.of.logs)+((2*aicetc$coefficients)*(aicetc$n/(aicetc$n - aicetc$coefficients-1))))
write.csv(aicetc, paste0("../output/maxent/p2/models/", timeslice, "/", modn, "AICforMaxentReg.csv"), row.names = FALSE)

@@

>>>>

```{r train and test inc AIC for reg val}
modelmonth <- cbind(paste0(mthna ,unique(prback$year)))
#modelmonth <- "jan2003"
aicetc <- read.csv("../data/env/AICcCalculations_p2.csv", header = TRUE) #pre-created csv file
betavallist <- read.csv("../data/env/AICcCalculations.csv", header = TRUE) #pre-created csv file
betavallist <- cbind(betavallist$regularization)
nfolds <- cbind(c(1:5))

for (year in 1:nrow(modelmonth)){

  timeslice <- modelmonth[year] #testing timeslice
  #timeslice <- modelmonth
  yr <- str_sub(timeslice, -4) #the year being used for testing
  train <- subset(prback, year == yr) #training data is all years but testing
  trainpa <- train$occurrence #creates a vector of presence/background ID (1/0)
  train  <- subset(train, select = -c(year, occurrence)) #don't need year and occurrence for modelling

  for (h in 1:nrow(betavallist)){
    betamval <-  betavallist[h]
    mod <- maxent(x = train, p = trainpa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("replicates=", nrow(nfolds)), "replicatetype=crossvalidate", paste0("betamultiplier=", betamval)), path = paste0("C:/Users/bunny/Documents/GitHub/Chapter_1/output/maxent/p2/models/", timeslice)) #here the full path is needed

@@

@@

    #count number of non-zero coefficients, excluding the final four rows
    lambdas_list <- list.files(paste0("../output/maxent/p2/models/", timeslice), pattern = "*lambdas", full.names = TRUE)
    samppred_list <- list.files(paste0("../output/maxent/p2/models/", timeslice), pattern = "*samplePredictions.csv", full.names = TRUE)

      for(f in 1:nrow(nfolds)){

      modn <- nfolds[f]-1
      lamf <- lambdas_list[f]
      sampf <- samppred_list[f]
      coefile <- read.csv(paste0(lamf), header = FALSE, stringsAsFactors=FALSE)
      coefile <- head(coefile, -4)
      coefile2 <- subset(coefile, V2 != 0)
      nocoeff <- nrow(coefile2)
      bick <- sum(as.numeric(coefile2[, 2])) #dodgy 
      bic <- (bick*log(sum(trainpa == 1)) - 2*11) #dodgy
      #natural log
      samplepred <- read.csv(paste0(sampf), header = TRUE)
      samplepred$nat_log <- log(samplepred$Raw.prediction)
      sumsamplepred <- sum(samplepred$nat_log)

@@

      notrainingpoints <- maxresults$X.Training.samples[f]
      testauc <- maxresults$Test.AUC[f]
      trainauc <- maxresults$Training.AUC[f]
      modno <- maxresults$Species[f]

      aicetc$train_AUC <- trainauc
      aicetc$test_AUC <- testauc
      aicetc$n <- notrainingpoints
      aicetc$coefficients <- nocoeff
      aicetc$sum.of.logs <- sumsamplepred

      aicetc$bic <- bic #not sure this is correct...
      aicetc$beta <- betamval
      aicetc$foldno <- modn

@@

    aicetc$AIC <- ((2*aicetc$coefficients)-(2*aicetc$sum.of.logs))
    aicetc$AICc <- (-2*(aicetc$sum.of.logs)+((2*aicetc$coefficients)*(aicetc$n/(aicetc$n - aicetc$coefficients-1))))
    write.csv(aicetc, paste0("../output/maxent/p2/models/", timeslice, "/", betamval, "_", modn, "_", "AICforMaxentReg.csv"), row.names = FALSE)
      }

  # aa <- paste0("fold", modn, "_", betamval, "_", "AICforMaxentReg.csv")
    aicperbetalst <- list.files(paste0("../output/maxent/p2/models/", timeslice, "/"), pattern = paste0("^", betamval, ".*AICforMaxentReg.csv"), full.names = TRUE)
    aicperbeta <- lapply(aicperbetalst, read.csv, header = TRUE)
    aicperbeta <- do.call(rbind, aicperbeta)
    aicbetamean <- colMeans(aicperbeta)
    aicperbeta <- rbind(aicperbeta, aicbetamean)
    aicperbeta$foldno[f+1] <- f+1
    write.csv(aicperbeta, paste0("../output/maxent/p2/models/", timeslice, "/aicperbetasum_", betamval, ".csv"), row.names = FALSE)

    betaavglst <- list.files(paste0("../output/maxent/p2/models/", timeslice, "/"), pattern = "aicperbetasum_.*csv", full.names = TRUE)
    bavg <- lapply(betaavglst, read.csv, header = TRUE)
    bavg <- do.call(rbind, bavg)
    bavg <- subset(bavg, foldno == f+1)
    write.csv(bavg, paste0("../output/maxent/p2/models/", timeslice, "/avgbetalst.csv"), row.names = FALSE)
  }

@@

#aicperfoldlst <- list.files(paste0("../output/maxent/p2/models/", timeslice), pattern = "fold4.*AICforMaxentReg.csv", full.names = TRUE)
#aicperfold <- lapply(aicperfoldlst, read.csv, header = TRUE)
#aicperfold <- do.call(rbind, aicperfold)
